---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# README

These are analysis operations and data for the study titled "Musical expertise predicts perceived tension in intervals better than psychoacoustics across North and South Indian listeners" by Imre Lahdelma, [Tuomas Eerola](https://tuomaseerola.github.io/), Nashra Ahmad, Martin Clayton, James Armitage, Budhaditya Bhattacharyya, and Niveshan Munsamy. Read this document at the [documentation site](https://tuomaseerola.github.io/interval_tension/).

## Load data

```{r load_data, warning = FALSE, message = FALSE}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(papaja)
options(dplyr.summarise.inform = FALSE)
df <- read.csv('data/indian_tension_ratings.csv')
knitr::kable(head(df),format = 'markdown')
```

## Describe participants

```{r describe_data, warning = FALSE, message = FALSE}
U<-as.character(unique(df$participant))
background <- NULL
for (k in 1:length(U)) {
  tmp <- dplyr::filter(df,participant==U[k])
  background<-rbind(background,tmp[1,])
}

SS <- summarise(group_by(background,Group),N=n(),ageM=mean(age,na.rm=TRUE),ageSE=sd(age,na.rm=TRUE)/sqrt(n()),expertiseM=mean(ExpertiseN,na.rm=TRUE),expertiseSE=sd(ExpertiseN,na.rm=TRUE)/sqrt(n()),female=sum(gender=='Female')/n()*100)

knitr::kable(SS,digits = 2,format = 'markdown')

m1 <- aov(ExpertiseN ~ Group, data=background)
a<-apa_print(summary(m1))
print(a$statistic)
print(TukeyHSD(m1))
```

## LMM analysis

### Are the Hindustani and Carnatic musicians tension ratings different from each other?

```{r LMM, warning = FALSE, message = FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
library(effectsize)

# Take only Carnatic and Hindustani Musicians
tmp<-dplyr::filter(df,Group=='Carnatic' | Group=='Hindustani')
tmp$Group<-factor(tmp$Group)
m1 <- lmer(tensionrating ~ Interval * Group + (1|participant), data=tmp)
print(anova(m1))
# No difference between the two expert groups nor interaction
```

### Differences between and within musicians and non-musicians?

```{r LMM2}
# Combine two musician groups and relabel everybody to musician/non-musician
df$Group2<-as.character(df$Group)
df$Group2[df$Group=='Carnatic']<-'Musicians'
df$Group2[df$Group=='Hindustani']<-'Musicians'
df$Group2[df$Group=='Indian - non-musicians']<-'Non-musicians'
df$Group2<-factor(df$Group2)

m1 <- lmer(tensionrating ~ Group2 * Interval + (1|participant),data = df)
print(anova(m1))
F_to_eta2(8.7738, 11, 757.09)  # effect size
F_to_eta2(10.4117, 11, 757.09) # effect size

print(pairs(emmeans(m1,~ Group2|Interval))) # this for each interval
```

#### Check discrimination of intervals within musicians

```{r LMM-musicians}
m1 <- lmer(tensionrating ~ Interval + (1|participant),data = dplyr::filter(df,Group2=='Musicians'))
print(anova(m1))
F_to_epsilon2(17.848, 11, 482.16) # 0.27 [0.21, 1.00]
emm <- emmeans(m1,specs = pairwise ~ Interval,pbkrtest.limit = 4345, type = "response")
emm$contrasts
# 34/66 comparison significant at p<.05 level
```

#### Check discrimination of intervals within non-musicians

```{r LMM-nonmusicians}
m1 <- lmer(tensionrating ~ Interval + (1|participant),data = dplyr::filter(df,Group2=='Non-musicians'))
anova(m1)
F_to_epsilon2(4.5092, 11, 275) # 0.12 [0.04, 1.00]
emm <- emmeans(m1,specs = pairwise ~ Interval,pbkrtest.limit = 4345, type = "response")

print(emm$contrasts)
# 7/66 comparison significant at p<.05 level
```


## Visualise tension ratings of both groups


```{r figure, fig.width=10.5,fig.height=7,warning=FALSE,message=FALSE,results='asis'}
# take the means and 95% CIs

S <- df %>%
  group_by(Interval,Group2) %>%
  summarize(n=n(),M=mean(tensionrating,na.rm = TRUE),sd=sd(tensionrating,na.rm = TRUE)) %>%
  mutate(se=sd/sqrt(n),LCI=M+qnorm(0.025)*se,UCI=M+qnorm(0.975)*se)

dval<-0.33
colnames(S)[2] <- 'Expertise'

fig_india <- ggplot(S,aes(Interval,M,color=Expertise,group=Expertise,linestyle=Expertise,shape=Expertise))+
  geom_point(position=position_dodge(dval),size=4)+
  geom_line(position=position_dodge(dval))+
  geom_errorbar(aes(min=LCI, max=UCI),width=.2,position=position_dodge(dval)) +
  scale_y_continuous(breaks = seq(1,7,by=1),limits = c(1,7),expand = c(0.005,0.005))+
  scale_color_brewer(type = "div",palette = 'Set2')+
  scale_shape_manual(values=c(19,17))+  
  ylab('Tension Rating (M ± 95% CI)')+
  theme_bw()+
  theme(legend.position="bottom")+
  # Annotations take the statistics from analyses above
  annotate("text",x = 1,y=5.87,label='italic(p)<.001',parse = TRUE,size=3)+
  annotate("text",x = 4,y=5.87,label='italic(p)<.05',parse = TRUE,size=3)+
  annotate("text",x = 5,y=5.87,label='italic(p)<.001',parse = TRUE,size=3)+
  annotate("text",x = 7,y=5.87,label='italic(p)<0.01',parse = TRUE,size=3)+
  annotate("text",x = 8,y=5.87,label='italic(p)<.05',parse = TRUE,size=3)+
  annotate("text",x = 9,y=5.87,label='italic(p)<.001',parse = TRUE,size=3)+
  annotate("text",x = 11,y=5.87,label='italic(p)<.05',parse = TRUE,size=3)+
  annotate("text",x = 12,y=5.87,label='italic(p)<.01',parse = TRUE,size=3)+
  annotate("text",x = 1,y=6.5,label='Between group comparisons',size=4,hjust=0)+
  annotate("text",x = 1,y=2.5,label='Within group comparisons',size=4,hjust=0)+
  annotate("text",x = 1.5,y=2.0,label='Musicians:~34/66~intervals~differ~at~italic(p)<.05',color=RColorBrewer::brewer.pal(3, 'Set2')[1],size=3,hjust=0,parse=TRUE)+
  annotate("text",x = 1.5,y=1.5,label='Non-musicians:~7/66~intervals~differ~at~italic(p)<.05',color=RColorBrewer::brewer.pal(3, 'Set2')[2],size=3,hjust=0,parse=TRUE)+
  ggtitle('Indian samples (N=71)')
print(fig_india)

```

#### Session information

R version and libraries.

```{r}
print(sessionInfo())
```

